---
output: html_document
---

The following examples are based on a data set from a 2002 field survey 
by T. Leyens which has been funded by the Deutsche Gesellschaft für 
Internationale Zusammenarbeit (GIZ). For more information, please refer to 
[this report](http://www2.gtz.de/dokumente/bib/02-5082.pdf). 
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(sp)
library(raster)
library(rgdal)
data_2007 <- readOGR(
  "D:/active/bis-fogo/school2015/data/field-campaign_2002/procd/IE_2007_pontos_esp.shp",
  "IE_2007_pontos_esp", stringsAsFactors = FALSE)
data <- as.data.frame(data_2007)
data$CYD_OBL <- as.numeric(data$CYD_OBL)
data_with_species <- which(rowSums((data[,14:89]>0))>0)
data <- data[data_with_species,]
data_alt_error <- which(data$ALT_GPS_M==0)
data <- data[-data_alt_error,]
data_species <- data[,14:89]
richness <- rowSums(data_species)
occupancy <- colSums(data_species>0)
species_considered <- which(occupancy>10)
```


#### Occurrence of species as a function of elevation
Again, we assume that elevation has some explanatory power. This time, the 
elevational distribution of selected species should be analyzed.

To visualize the results, a binary plot (i.e. presence/absence) is computed
for each species and a binominal generalized linear model is fitted to this
distribution using elevation (and it's square) as explanatory variable.

In order to compute this for 8 species (defined in variable "species_considered"),
a for-loop is used. Within each loop, a plot for the respective species is
generated and a GLM model is fitted afterwards. To get the elevational function
of species occurrence, the fitted model is used to predict the respective
occurrence at height intervals of 1 meter between the minimum and maximum 
elevation for which species data has been recorded.
```{r}
# Set graphical output to 3 by 3 columns/rows
user<-par(no.readonly=T)
par(mar=c(4,4,1,1),las=1,cex=0.9,mfcol=c(3,3))

# Store elevation in variable x (just for technical reasons of the prediction
# function)
x <- data$ALT_GPS_M
predict_x <- min(x):max(x)

# Compute elevational distribution for each species stored in variable 
# species_considered
for (i in species_considered)
{ 
  # Compute plot
  plot(data$ALT_GPS_M, data_species[,i]>0,
       xlab="Elevation [m]",ylab=paste("Pres/Abs",names(data_species)[i]),
       pch=21,bg="grey")
  
  # Train model
  model <- glm((data_species[,i]>0) ~ x + I(x^2),family="binomial")
  
  # Use model to predict species elevational distribution
  pred <- predict(model, newdata = data.frame(x = predict_x))
  
  # Add prediction results to plots
  lines(predict_x, exp(pred) / (1 + exp(pred)))
}
par(user)
```
