---
output: html_document
---

The following examples are based on a data set from a 2002 field survey 
by T. Leyens which has been funded by the Deutsche Gesellschaft für 
Internationale Zusammenarbeit (GIZ). For more information, please refer to 
[this report](http://www2.gtz.de/dokumente/bib/02-5082.pdf). 
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(sp)
library(raster)
library(rgdal)
data_2007 <- readOGR(
  "D:/active/bis-fogo/school2015/data/field-campaign_2002/procd/IE_2007_pontos_esp.shp",
  "IE_2007_pontos_esp", stringsAsFactors = FALSE)
data <- as.data.frame(data_2007)
data$CYD_OBL <- as.numeric(data$CYD_OBL)
data_with_species <- which(rowSums((data[,14:89]>0))>0)
data <- data[data_with_species,]
data_alt_error <- which(data$ALT_GPS_M==0)
data <- data[-data_alt_error,]
data_species <- data[,14:89]
richness <- rowSums(data_species)
occupancy <- colSums(data_species>0)
species_considered <- which(occupancy>10)
```


Regarding libraries, the following packages are necessary.
```{r, warning=FALSE}
library(vegan)
```

#### Multivariate data mining
Ordination analysis is a multivariate data mining method which reduces the 
dimensionality of an ecological data set and re-arranges the reduced data values 
onto a two dimensional space in such a way that potential relationship patterns 
between the individual ecological information becomes as apparent as possible. 
Therefore, common data reduction and transformation techniques like principal 
component analysis, multi-dimensional scaling or correspondence analysis is used. 

 Maybe the most commonly applied technique in ordination analysis is the 
 correspondence analysis. It is heavily used to identify and delineate 
 communities of species (i.e. groups of species which frequently occur together 
 at one location). The correspondence analysis uses a Chi-square distance matrix 
 as basis for its eigenalanysis. Therefore, the observed vs. expected values of 
 each species at each location are computed following the commonly used 
 Chi-square formula:

`(observed value - expected value)2 / expected value`

In addition, environmental factors can be considered as so called constraining 
variables. 


```{r}
ordination<-cca(data_species ~ GRAU_EROS + SOLO + CAT_USO + DECL_GR + EXP_GR +
                  MAT_ORG + ALT_GPS_M + GRAU_UTIL, data = data)
plot(ordination)
```


In order to get an idea on the significance of the constraining variables, a simple
analysis of variance (ANOVA) can be calculated using the ordination object.
```{r}
anova(ordination, by = "terms", permu = 999)
```
